### 2024-12-03 paper/sections/abstract.tex
paper: add template

\documentclass[../main.tex]{subfiles}
\begin{document}
Your abstract text goes here. Just a few facts. Whet our appetites.
Not more than 200 words, if possible, and preferably closer to 150.
\end{document}