### 2024-12-03 paper/sections/design.tex

\documentclass[../main.tex]{subfiles}
\begin{document}
Design
\end{document}