### 2024-12-03 paper/sections/introduction.tex

\documentclass[../main.tex]{subfiles}
\begin{document}
A paragraph of text goes here. Lots of text. Plenty of interesting
text. Text text text text text text text text text text text text text
text text text text text text text text text text text text text text
text text text text text text text text text text text text text text
text text text text text text text.
More fascinating text. Features galore, plethora of promises.
\end{document}