### 2024-12-03 paper/sections/relatedwork.tex

\documentclass[../main.tex]{subfiles}
\begin{document}
Related work
\end{document}