### 2024-12-06 paper/sections/relatedwork.tex

\documentclass[../main.tex]{subfiles}
\begin{document}
CCP decoupled data gathering and CCA computation. A datapath module gathers information regarding a TCP flow and sends it to user space via various methods (such as netlink sockets and unix pipes), while a user space program listens to the given data and sends back any updates to be made to the TCP connection state. Platform/OS independence
\end{document}