### 2024-12-10 paper/sections/conclusion.tex

\documentclass[../main.tex]{subfiles}
\begin{document}
Conclusion
\end{document}