### 2024-12-12 paper/sections/conclusion.tex

\documentclass[../main.tex]{subfiles}
\begin{document}
In this paper, we have evaluated the various attributes of eBPF as a platform for CCA implementation. First, through our behavior evaluations, we have shown that CCAs implemented through alterantive platforms (eBPF, CCP, eBPFCCP) do achieve behavioral correctness. We have also shown that eBPF is a viable platform for developing CCAs natively on, with negligible performance overhead compared to the kernel native implementation. Our implementation of eBPFCCP had a higher performance overhead, but was still able to achieve a significant percentage of the full kernel implementation value. We believe this artifact can be useful as an alternative CCP datapath that provides safety guarantees. Overall, the data demonstrates that using eBPF-related CCA platforms not only provides many benefits to development speed and safety, but can also be used as a long-term deployment option outside of research stages. 
\end{document}