### 2024-12-12 paper/sections/conclusion.tex

\documentclass[../main.tex]{subfiles}
\begin{document}
In this paper, we introduced an eBPF-based platform called eBPFCCP, which enables the implementation of user space CCAs. We then compared eBPFCCP with the native Linux kernel TCP stack, a pure eBPF implementation, and CCP. Our evaluations focused on the behavioral correctness and performance impact of these implementations. Through our behavior evaluations, we demonstrated that CCAs implemented using alternative platforms achieve the desired behavior. Additionally, we showed that eBPF is a suitable platform for developing CCAs natively, with negligible performance overhead compared to the kernel’s native implementation. While our implementation of eBPFCCP had a slightly higher performance overhead, it still achieved a substantial percentage of the full kernel implementation’s value. We believe that eBPFCCP can be valuable as an alternative CCP datapath that offers safety guarantees. Overall, the data presented here demonstrates that using eBPF-related CCA platforms not only provides numerous benefits, including enhanced development speed and safety, but also has the potential for long-term deployment beyond research stages. 
\end{document}