### 2024-12-12 paper/sections/evaluation.tex

\documentclass[../main.tex]{subfiles}
\begin{document}

\subsection{Setup and Method}
An alternative implementation of an existing CCA can be evaluated using two metrics.
We evaluate eBPFCCP using two metrics:
\begin{enumerate}
    \item {\bf Behavior}: how accurate is eBPFCCP in replicating original CCA behavior?
    \item {\bf Performance}: what is the performance overhead of eBPFCCP?
\end{enumerate}

We included two sets of benchmarks to measure this: (1) a behavioral benchmark aimed to verify the "correctness" of different implementations (TCP Cubic should behave similarly no matter where it is implemented, as indicated by metrics such as bitrate/congestion window size over time), and (2) a performance benchmark aimed to measure the CPU overhead of different implementation approaches. A CCA implementation with smaller CPU overhead would logically lead to a higher maximum bandwidth in an ideal, infinite-speed connection where the CPU running the CCA is the bottleneck. \\
Our benchmarks utilize iperf3 \cite{iperf}. For each trial, we loop through every CCA we are testing and run iperf3 for $15$ seconds. At the same time, we run a Python script that loops infinitely while incrementing a variable. Once the $15$ seconds are up, the script then outputs the final number into a file. A higher output number suggests that the script went through more CPU cycles during the trial, indicating better CPU performance. We a ran a total of $20$ trials. \\
After all the trials, we average the behavioral metric across the trials and plot the average for each CCA through time as the behavioral benchmark. We then plot the CPU performances for each CCA across trials on a boxplot. \\
To make sure we are correctly benchmarking the performance impact of CCAs rather than miscellneous confounding factors, we ran the benchmark on an out-of-the-box Debian image on a VM. To ensure the CPU benchmark and the CCA are running on the same core, our VM is set up to use only one core. We also decided to run our benchmarks on various implementations of TCP Cubic due to its wide availability across platforms.

\subsection{Results}

\begin{figure}[t]
    \centering
        \includegraphics[width=\columnwidth]{img/cca-behavior-cwnd}
        \caption{snd-cwnd behavior of TCP Cubic across different platforms of implementation. Averaged over 20 trials.
        }\label{fig:cca-behavior-cwnd}
    \end{figure}

\begin{figure}[t]
    \centering
        \includegraphics[width=\columnwidth]{img/cca-behavior-bitrate}
        \caption{Bitrate behavior of TCP Cubic across different platforms of implementation. Averaged over 20 trials.
        }\label{fig:cca-behavior-bitrate}
    \end{figure}

\begin{figure}[t]
    \centering
        \includegraphics[width=\columnwidth]{img/cca-behavior-rtt}
        \caption{RTT behavior of TCP Cubic across different platforms of implementation. Averaged over 20 trials.
        }\label{fig:cca-behavior-rtt}
    \end{figure}

\begin{figure}[t]
    \centering
        \includegraphics[width=\columnwidth]{img/cca-performance}
        \caption{Standard box plot of TCP Cubic CPU performance across different platforms of implementation. Higher is better. Box extends from 25th to 75th percentile. Whiskers extend to the farthest data point within 1.5 IQR from the box. Data gathered over 20 trials.
        }\label{fig:cca-performance}
    \end{figure}

The results of the behavioral benchmarks are shown in Figures~\ref{fig:cca-behavior-cwnd},~\ref{fig:cca-behavior-bitrate}, and~\ref{fig:cca-behavior-rtt}. From the graphs, we can see that the behavior of the same Congestion Control Algorithm remains similar across congestion window size, bitrate, and RTT. The results of the CPU benchmarks are shown in Figure \ref{fig:cca-performance}. The eBPF implementation of TCP Cubic achieved similar CPU performance compared to the built-in Linux kernel networking stack implementation. CCP-Cubic from the CCP Project \cite{generic-cong-avoid} combined with the CCP datapath kernel module and netlink slightly lower performance, where its median is at x percent of the full kernel implementation value. Our implementation of eBPFCCP ranked lower on the performance chart, but was still able to achieve y percent of the full kernel implementation value. Due to the relative comparability of CPU performance between the kernel and eBPF implementations of TCP Cubicm, we theorize the gap between eBPFCCP and kernel module CCP was due to our choice of using unix pipes rather than netlink as a channel for communicating between the eBPFCCP and the user space program. \\
From these results, we can conclude that eBPF is a viable platform for the development CCAs. The minor performance overhead when compared to the kernel native implementation suggests that eBPF-based CCAs can be used as a long-term deployment option outside of research and development stages. With regards to CCP, while the performance overhead of eBPFCCP is non-negligible, it should be noted that the benchmark is performed on a single core VM. In a real-world scenario, the performance overhead of eBPFCCP is likely to be negligible for all but the most performance-sensitive industry networking applications. The safety guarantees provided by eBPFCCP would provide a good intermediate step for users to adopt and experiment with a new CCA before it is natively implemented either directly into the kernel or as an eBPF program. 
\end{document}